An equivalence of compositional frameworks

### 2019-05-30 EquivCompFrameworks_20180911kc.tex
why not fibrations in appendix as well?

\begin{itemize}
\item {\chris Discuss notation for Decorated and Structured cospan double categories one last time? :)}
\item {\chris Notation between \cref{Equiv} and \cref{thm:decoratedcospans} right now does not match! Need to fix. Slightly prefer a,b,c for objects of A, then $a'$ and $b'$ for other objects, and x an object of the category F(c) namely the decoration. The previously used `d' is slightly confusing around all these a's, b's and c's that are in a different category. Also I prefer $a$ and $b$ rather than $a_1$ and $a_2$.}
%\item {\chris There might be a way to completely hide the details of version 'fibres have, reindexing preserve' if that's what we want to do.}
\end{itemize}

@@

{\chris moved here preliminary}
In this section, we recall the double category of \emph{structured cospans}, a formalism introduced to the first two authors. One of the main goals of this paper is to provide a double equivalence between this double category and one of \emph{decorated cospans}, described in detail in the subsequent section.

{\chris Wait, why do you need to consider them as monoidal in the assumptions of the theorem?}

\begin{thm}\label{SC}
Given a category $\mathsf{X}$ with finite colimits and a category $\mathsf{A}$ with finite coproducts and a finite coproduct preserving functor $L \colon \mathsf{A} \to \mathsf{X}$ with $\mathsf{A}$ and $\mathsf{X}$ regarded as cocartesian monoidal categories, there exists a symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$ which has:
\begin{enumerate}
\item{objects given by objects of $\mathsf{A}$,}
\item{vertical 1-morphisms given by morphisms of $\mathsf{A}$,}
\item{horizontal 1-cells given by cospans of $\mathsf{X}$ of the form:
\begin{equation}\label{eq:structuredcospan}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(c)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B);
\end{tikzpicture}
\end{equation}
and}
\item{2-morphisms given by maps of cospans of $\mathsf{X}$ of the form:
\begin{equation}\label{eq:2cellsStrCsp}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(c_1)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(c_2)$};
\node (A') at (0,-1) {$L(c_1^\prime)$};
\node (B') at (1,-1) {$x^\prime$};
\node (C') at (2,-1) {$L(c_2^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A') edge node[above]{$$} (B')
(C') edge node[above]{$$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [left]{$L(g)$} (C');
\end{tikzpicture}
\end{equation}
Composition of horizontal 1-cells and 2-morphisms is given by pushouts in $\mathsf{X}$ and tensoring of objects is under binary coproducts in $\mathsf{A}$.
}
\end{enumerate}
\end{thm}
\begin{proof}
See the first two authors' work on structured cospans \cite{BC2}.
\end{proof}
{\chris Below, I commented out a nice short recalled version of decorated cospans, since now their definition is later. Use if we want to recall at some point later}
%In the previous section, we started with a symmetric lax monoidal pseudofunctor $F \colon \mathsf{A} \to \mathbf{Cat}$ and created a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ which has:
%\begin{enumerate}
%\item{objects of $\mathsf{A}$ as objects,}
%\item{morphisms of $\mathsf{A}$ as vertical 1-morphisms,}
%\item{horizontal 1-cells are given by $F$-decorated cospans, which are pairs:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$c_1$};
%\node (B) at (1,0) {$c$};
%\node (C) at (2,0) {$c_2$};
%\node (D) at (3,0) {$x \in F(c)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$i$} (B)
%(C) edge node[above]{$o$} (B);
%\end{tikzpicture}
%\]
%and}
%\item{2-morphisms are given by maps of cospans in $\mathsf{A}$:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$c_1$};
%\node (B) at (1,0) {$c$};
%\node (C) at (2,0) {$c_2$};
%\node (A') at (0,-1) {$c_1^\prime$};
%\node (B') at (1,-1) {$c^\prime$};
%\node (C') at (2,-1) {$c_2^\prime$};
%\node (D) at (3,0) {$x \in F(c)$};
%\node (D') at (3,-1) {$x^\prime \in F(c^\prime)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$i$} (B)
%(C) edge node[above]{$o$} (B)
%(A') edge node[above]{$i^\prime$} (B')
%(C') edge node[above]{$o^\prime$} (B')
%(A) edge node [left]{$f$} (A')
%(B) edge node [left]{$h$} (B')
%(C) edge node [left]{$g$} (C');
%\end{tikzpicture}
%\]
%together with a morphism $\iota \colon F(h)(d) \to d^\prime$ in $F(c^\prime)$.}
%\end{enumerate}

{\chris here ends old}

In this section we prove that the two different double categorical network frameworks, namely structured cospans and decorated cospans described respectively in \cref{DecCospansDoubleCat,sec:structuredcospans}, are equivalent under certain assumptions. This is established by \cref{Equiv}, which relies on conditions relevant to the existence of left adjoints for opfibrations. We begin by sketching this underlying framework, which is worked out in detail in \cite{MV} and \cite{CV}.
%In this section we investigate the necessary conditions to obtain a category $\mathrm{X}$ and a left adjoint $L \colon \mathrm{A} \to \mathrm{X}$ from a category $\mathrm{A}$ with finite colimits and a symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \mathbf{Cat}$. The main tool that allows us to do this will be the Grothendieck construction which we recall:
First of all, we recall some basics regarding opfibrations and their relation to pseudofunctors into $\bicat{Cat}$ via the Grothendieck construction. Standard material on the theory of fibrations can be found for example in \cite{Herm}, \cite{Borc}.

Let $\OpFib(\A)$ denote the 2-subcategory of the slice 2-category $\mathbf{Cat}/ \A$ of opfibrations over $\A$, functors that preserve cocartesian liftings and natural transformations with vertical components. In fact, there is a 2-equivalence between opfibrations and pseudofunctors, see Section ?, induced by the so-called \emph{Grothendieck construction}.
%{\chris I don't mind this presentation of the Grothendieck construction which I admit not having seeing before. If we are to keep, we should put a reference. Otherwise we could stay more low level about this and just briefly describe what's going on.}
%\begin{defn}
%Let $\mathbf{Cat}_{\textnormal{lax},\star}$ denote the \define{2-category of lax-pointed categories} which has:
%\begin{enumerate}
%\item{objects as pairs $(\mathrm{A},c)$ where $\mathrm{A}$ is a category and $c$ is an object of $\mathrm{A}$, and}
%\item{a morphism from $(\mathrm{A},c)$ to $(\mathrm{X},y)$ is a pair $(F,f)$ where $F \colon \mathrm{A} \to \mathrm{X}$ is a functor and $f \colon F(c) \to y$ is a morphism in $\mathrm{X}$.}
%\end{enumerate}
%\end{defn}
%\begin{defn}
%Given a pseudofunctor $F \colon \mathrm{A} \to \mathbf{Cat}$, the \define{Grothendieck construction of} $F$ is given by the `strict 2-pullback' of the following cospan:
% \[
%\begin{tikzpicture}[scale=1.5]
%\node (D) at (1,1) {$\inta F$};
%\node (A) at (0,0) {$\mathrm{A}$};
%\node (B) at (1,-1) {$\mathbf{Cat}$};
%\node (C) at (2,0) {$\mathbf{Cat}_{\textnormal{lax},\star}$};
%\path[->,font=\scriptsize,>=angle 90]
%(D) edge[dashed] node[above]{$p$} (A)
%(D) edge[dashed] node[above]{$q$} (C)
%(A) edge node[below]{$F$} (B)
%(C) edge node[below]{$P$} (B);
%\end{tikzpicture}
%\]
%where $P \colon \mathbf{Cat}_{\textnormal{lax},\star} \to \mathbf{Cat}$ is the forgetful functor. 
\begin{defn}\label{def:GrothCat}
For any pseudofunctor $F\colon\A\to\bicat{Cat}$ where $\A$ is a category viewed as a 2-category with trivial 2-cells, the \textbf{Grothendieck category}
$\inta F$ has
\begin{itemize}
\item objects pairs $(a, x \in F(a))$ and
\item a morphism from $(a, x \in F(a))$ to $(b, y\in F(b))$ is a pair $(f \colon a \to b,\delta \colon F(f)(x) \to y)\in\A\times F(b)$.
\end{itemize}
{\chris Might need to write the composition rule? Depending on the final proof of Big Theorem.}
This is an opfibred category over $\A$ via the obvious forgetful functor, with fibre categories $(\inta F)_a=F(a)$ and reindexing functors $f_!=F(f)$.
\end{defn}
%This can also be thought of as a morphism and a 2-morphism:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$\star$};
%\node (B) at (1,0.5) {$F(c)$};
%\node (C) at (1,-0.5) {$F(c^\prime)$};
%\node (D) at (2,0.5) {$c$};
%\node (E) at (2,-0.5) {$c^\prime$};
%\node (F) at (0.65,0) {$\Swarrow \alpha$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$d$} (B)
%(B) edge node[right]{$F(f)$} (C)
%(A) edge node[below]{$d^\prime$} (C)
%(D) edge node[left] {$f$} (E);
%\end{tikzpicture}
%\]
The above in fact provides the one direction of the following well-known equivalence.
%gives rise to a 2-functor $[\A,\bicat{Cat}]_\pse\to \mathbf{Cat}/ \A$ which factors through the embedding $\OpFib(\A) \hookrightarrow \bicat{Cat} / \A$ as per the following theorem.
%: $$\inta F \colon [ \mathrm{A}, \mathbf{Cat} ] \to \textnormal{Opfib}(\mathrm{A}) \hookrightarrow \mathbf{Cat} / \mathrm{A}.$$

@@

\begin{thm}\label{Equiv}
Suppose $\ca{A}$ is a finitely cocomplete category and $(F,\phi,\phi_0)\colon(\ca{A},+,0) \to (\bicat{Cat},\times,\mathbf{1})$ is a symmetric lax monoidal pseudofunctor. %such that as an ordinary pseudofunctor, it factors through $\bicat{fcocCat}$. 
Then the symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ of decorated cospans (\cref{thm:decoratedcospans}) is equivalent to the symmetric monoidal double category $\scsp{L_F}{\inta F}$ of structured cospans (\cref{SC}), where $L_F \colon \mathsf{A} \to \inta{F}$ is the left adjoint of the induced Grothendieck opfibration $\U_F\colon \inta{F} \to \mathsf{A}$ %by \cref{thm:Grothendieck}, 
under the assumptions of \cref{cor:fcocMonGroth}.
\end{thm}

\begin{proof}
Recall that the double category of decorated cospans $\dcsp{F}$ has objects and vertical 1-morphisms those of $\A$, horizontal 1-cells $a\tickar b$ are cospans $\cspn{a}{c}{b}{}{}$ in $\A$ along with $x\in F(c)$, and 2-morphisms are cospan maps $k\colon c\to c'$ together with a morphism $(Fk)(x)\to x'$, see \cref{eq:FCsp2morph}.

@@

>>>>

\begin{itemize}
\item {\chris Would a title like `An equivalence of compositional frameworks for networks' or `An equivalence of compositional network frameworks' be more informative? Otherwise, compositional framework on its own doesn't say much!}
\item {\chris Discuss notation for Decorated and Structured cospan double categories one last time? :)}
\item {\chris Notation between \cref{Equiv} and \cref{thm:decoratedcospans} right now does not match! Need to fix. Slightly prefer a,b,c for objects of A, then $a'$ and $b'$ for other objects, and x an object of the category F(c) namely the decoration. The previously used `d' is slightly confusing around all these a's, b's and c's that are in a different category. Also I prefer $a$ and $b$ rather than $a_1$ and $a_2$.}
%\item {\chris There might be a way to completely hide the details of version 'fibres have, reindexing preserve' if that's what we want to do.}
\end{itemize}

@@

In \cite{BC2}, the first two authors introduces the symmetric monoidal double category of \emph{structured cospans} as a formalism to capture open networks. One of the main goals of this paper is to provide a monoidal double equivalence between this double category and that of \emph{decorated cospans}, described in detail in \cref{DecCospansDoubleCat}. Below we recall the double category structured cospans form.

{\chris Wait, why do you need to consider them as monoidal in the assumptions of the theorem?}
\begin{thm}\label{SC}
Given a category $\mathsf{X}$ with finite colimits and a category $\mathsf{A}$ with finite coproducts and a finite coproduct preserving functor $L \colon \mathsf{A} \to \mathsf{X}$ with $\mathsf{A}$ and $\mathsf{X}$ regarded as cocartesian monoidal categories, there exists a symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$ which has:
\begin{itemize}
\item{objects given by objects of $\mathsf{A}$,}
\item{vertical 1-morphisms given by morphisms of $\mathsf{A}$,}
\item{horizontal 1-cells $a\tickar b$ given by cospans of $\mathsf{X}$ of the form:
\begin{equation}\label{eq:structuredcospan}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B);
\end{tikzpicture}
\end{equation}}
\item 2-morphisms given by maps of cospans of $\mathsf{X}$ of the form:
\begin{equation}\label{eq:2cellsStrCsp}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\node (A') at (0,-1) {$L(a')$};
\node (B') at (1,-1) {$x'$};
\node (C') at (2,-1) {$L(b')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A') edge node[above]{$$} (B')
(C') edge node[above]{$$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [right]{$L(g)$} (C');
\end{tikzpicture}
\end{equation}
\end{itemize}
Composition of horizontal 1-cells and 2-morphisms is given by pushouts in $\mathsf{X}$ and tensoring of objects is under binary coproducts in $\mathsf{A}$.
\end{thm}
\begin{proof}
See the first two authors' work on structured cospans \cite{BC2}.
\end{proof}

The desired equivalence between these two different double categorical network frameworks, namely structured cospans and decorated cospans,
is established by the following theorem. We denote by $\bicat{fcocCat}$ the 2-category of finitely cocomplete categories and finitely cocontinuous functors.

\begin{thm}\label{Equiv}
Suppose $\ca{A}$ is a finitely cocomplete category and $(F,\phi,\phi_0)\colon(\ca{A},+,0) \to (\bicat{Cat},\times,\mathbf{1})$ is a symmetric lax monoidal pseudofunctor, which factors through $\bicat{fcocCat}$ as an ordinary pseudofunctor. 
Then the symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ of decorated cospans (\cref{thm:decoratedcospans}) is equivalent to the symmetric monoidal double category $\scsp{L_F}{\inta F}$ of structured cospans (\cref{SC}), where $L_F \colon \mathsf{A} \to \inta{F}$ is the left adjoint of the induced Grothendieck opfibration $\U_F\colon \inta{F} \to \mathsf{A}$. %by \cref{thm:Grothendieck}, 
%under the assumptions of \cref{cor:fcocMonGroth}.
\end{thm}

The conditions of the above theorem relate to the existence of left adjoints for opfibrations; we first sketch this underlying framework, and then we proceed to the proof of the theorem.

@@

\begin{proof}[Proof of \cref{Equiv}]
Recall that the double category of decorated cospans $\dcsp{F}$ has objects and vertical 1-morphisms those of $\A$, horizontal 1-cells $a\tickar b$ are cospans $\cspn{a}{c}{b}{}{}$ in $\A$ along with $x\in F(c)$, and 2-morphisms are cospan maps $k\colon c\to c'$ together with a morphism $(Fk)(x)\to x'$, see \cref{eq:FCsp2morph}.

@@

\subsection{Fibrations}\label{sec:fibrations}
We now recall some basics regarding opfibrations and their relation to pseudofunctors into $\bicat{Cat}$ via the Grothendieck construction. Standard material on the theory of fibrations can be found for example in \cite{Herm}, \cite{Borc}.

Let $\OpFib(\A)$ denote the 2-subcategory of the slice 2-category $\mathbf{Cat}/ \A$ of opfibrations over $\A$, functors that preserve cocartesian liftings and natural transformations with vertical components. In fact, there is a 2-equivalence between opfibrations and pseudofunctors induced by the so-called \emph{Grothendieck construction}.
%{\chris I don't mind this presentation of the Grothendieck construction which I admit not having seeing before. If we are to keep, we should put a reference. Otherwise we could stay more low level about this and just briefly describe what's going on.}
%\begin{defn}
%Let $\mathbf{Cat}_{\textnormal{lax},\star}$ denote the \define{2-category of lax-pointed categories} which has:
%\begin{enumerate}
%\item{objects as pairs $(\mathrm{A},c)$ where $\mathrm{A}$ is a category and $c$ is an object of $\mathrm{A}$, and}
%\item{a morphism from $(\mathrm{A},c)$ to $(\mathrm{X},y)$ is a pair $(F,f)$ where $F \colon \mathrm{A} \to \mathrm{X}$ is a functor and $f \colon F(c) \to y$ is a morphism in $\mathrm{X}$.}
%\end{enumerate}
%\end{defn}
%\begin{defn}
%Given a pseudofunctor $F \colon \mathrm{A} \to \mathbf{Cat}$, the \define{Grothendieck construction of} $F$ is given by the `strict 2-pullback' of the following cospan:
% \[
%\begin{tikzpicture}[scale=1.5]
%\node (D) at (1,1) {$\inta F$};
%\node (A) at (0,0) {$\mathrm{A}$};
%\node (B) at (1,-1) {$\mathbf{Cat}$};
%\node (C) at (2,0) {$\mathbf{Cat}_{\textnormal{lax},\star}$};
%\path[->,font=\scriptsize,>=angle 90]
%(D) edge[dashed] node[above]{$p$} (A)
%(D) edge[dashed] node[above]{$q$} (C)
%(A) edge node[below]{$F$} (B)
%(C) edge node[below]{$P$} (B);
%\end{tikzpicture}
%\]
%where $P \colon \mathbf{Cat}_{\textnormal{lax},\star} \to \mathbf{Cat}$ is the forgetful functor. 
\begin{defn}\label{def:GrothCat}
For any pseudofunctor $F\colon\A\to\bicat{Cat}$ where $\A$ is a category viewed as a 2-category with trivial 2-cells, the \textbf{Grothendieck category}
$\inta F$ has
\begin{itemize}
\item objects pairs $(a, x \in F(a))$ and
\item a morphism from $(a, x \in F(a))$ to $(b, y\in F(b))$ is a pair $(f \colon a \to b,\delta \colon F(f)(x) \to y)\in\A\times F(b)$.
\end{itemize}
{\chris Might need to write the composition rule? Depending on the final proof of Big Theorem.}
This is an opfibred category over $\A$ via the obvious forgetful functor, with fibre categories $(\inta F)_a=F(a)$ and reindexing functors $f_!=F(f)$.
\end{defn}
%This can also be thought of as a morphism and a 2-morphism:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$\star$};
%\node (B) at (1,0.5) {$F(c)$};
%\node (C) at (1,-0.5) {$F(c^\prime)$};
%\node (D) at (2,0.5) {$c$};
%\node (E) at (2,-0.5) {$c^\prime$};
%\node (F) at (0.65,0) {$\Swarrow \alpha$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$d$} (B)
%(B) edge node[right]{$F(f)$} (C)
%(A) edge node[below]{$d^\prime$} (C)
%(D) edge node[left] {$f$} (E);
%\end{tikzpicture}
%\]
The above in fact provides the one direction of the following well-known equivalence.
%gives rise to a 2-functor $[\A,\bicat{Cat}]_\pse\to \mathbf{Cat}/ \A$ which factors through the embedding $\OpFib(\A) \hookrightarrow \bicat{Cat} / \A$ as per the following theorem.
%: $$\inta F \colon [ \mathrm{A}, \mathbf{Cat} ] \to \textnormal{Opfib}(\mathrm{A}) \hookrightarrow \mathbf{Cat} / \mathrm{A}.$$
